---
title: 'Course notes: Day 2'
output:
  html_document:
    df_print: paged
    toc: true
    toc_depth: 4
    toc_float: true
    collapse : false
---

# Functions

## Create custom functions {.tabset}

Let's create a function that feeds an arbitrary number of dataframes with a given dataset. The two arguments will be:

- *data*, the input data;

- *times*, the number of assignments that are to be done (with a default of 5).

As R loads every assignment in memory, say we also want to include some safeguard such that the function generates a warning if one tries to assign the dataset more than 9 times and doesn't execute at all if more than 99 times. This can be done including the functions `warning()` and `stop()` within an if statement.

```{r, echo=T, results='hide',message=FALSE}
install.packages("pacman",repos = "http://cran.us.r-project.org")
pacman::p_load("here","tidyverse","readr","lubridate")
load("Data/ncbikecrash.Rdata") 
```


```{r}
# create an empty list
dupDF <- function(data,times=5){ # default value for times is 5
list.loop <- list() # create an empty list that will store the outputs
if(times>=10){
  if(times>=100){
    stop("Can't duplicate that many times")
  } else warning("Check you haven't screwed your machine's memory")
} else{
  for (i in 1:times){
     list.loop[[i]] = data
     names(list.loop)[i] <- paste("df_",i)
    }
    return(list.loop)
  }

}
```

## Store and re-use custom functions

Now that we've created `dupDF()`, let's save its script to a new `.R` file in a folder of the project called "Functions". The custom function can then be loaded and used from any other script by specifying its path as argument of the `source()` function.

```{r, message=F}
source(here::here("Functions","dupDF.R"))
```
Also, as for other functions of the tidyverse, `x %>% f(y)` is equivalent to `f(x,y)`, with x being the argument *data* and y the argument *times*. In other words, the data argument can be passed as follows:
```{r}
ncbikecrash %>%
  filter(region=="Piedmont") %>%
  distinct(region,city) %>%
  dupDF(10)
```

# Data formatting {.tabset}

```{r loadingData, message=FALSE}
apps_ratings <- read_csv(here::here("Data","googleplaystore.csv"))
user_reviews <- read_csv(here::here("Data","googleplaystore_user_reviews.csv"))
```

After loading the Google Play Store Apps, there are several operations Radia undertakes to make the dataset ready for analysis.

## Clean numbers

An issue is that the variable Size has values expressed in thousands (k) and in millions (M). We have to remove the k and M symbols and express everything in the same unit. Let's also remove the dollar sign in Price. The variable Install should also be cleaned from its end "+" and the commas it contains. Finally, we use `as.numeric()` to make sure our cleaned variables are doubles.
```{r, warning=FALSE}
apps_ratings_clean <- apps_ratings %>%
  filter(grepl("M", Size) | grepl("k", Size) ) %>%
  mutate(
    Size_cleaned = ifelse(grepl("M", Size),as.numeric(gsub("M", "",Size))*1000000,Size),
    Size_cleaned = ifelse(grepl("k", Size_cleaned),as.numeric(gsub("k", "",Size))*1000,Size_cleaned),
    Size_cleaned = as.numeric(Size_cleaned),
    Price_cleaned = as.numeric(gsub("\\$", "", as.character(Price))),
    Installs_cleaned = gsub("\\+", "", as.character(Installs)), # replace the end "+" by ""
    Installs_cleaned = as.numeric(gsub(",", "", Installs_cleaned)),
    Rating = as.numeric(Rating)
  )  %>%
  select(Size,Size_cleaned,Installs,Installs_cleaned, Price, Price_cleaned,`Last Updated`,Rating,Category,Type,App,Reviews)

options(scipen=999) ##Remove scientific notation

apps_ratings_clean %>% select(Size,Size_cleaned,Installs,Installs_cleaned, Price, Price_cleaned)
```

## Clean dates

The first step here is to convert the string variable `Last Updated` into a date format that R can understand.

```{r}
apps_ratings_clean <- apps_ratings_clean %>%
mutate(
    LastUpdated = as.POSIXct(apps_ratings_clean$`Last Updated`,format = "%d-%b-%y",tz=Sys.timezone())
)

apps_ratings_clean %>% select(`Last Updated`,LastUpdated)
```

From then on we can for example use functions from the lubridate package to play with the date's components. 

```{r, warning=FALSE}
apps_ratings_clean %>%
  mutate(
    year = lubridate::year(LastUpdated),
    month = lubridate::month(LastUpdated),
    yearQ = lubridate::yq(LastUpdated)
  ) %>%
  select(LastUpdated,year,month)
```


## Remove duplicates
In base R:
```{r}
apps_ratings_clean=apps_ratings_clean[!duplicated(apps_ratings_clean), ]
```
Using the tidyverse:
```{r}
apps_ratings_clean <- apps_ratings_clean %>% filter(!duplicated(apps_ratings_clean))
```

# Data viz with ggplot2

The first step is to specify the data and the axis, this is what's done in `ggplot(data= cars, aes(x = speed ,y= dist))`. Then you can add `geoms` to your visualization. `Geoms` behave like layers that you add to your graph. Labels can also be
```{r}
ggplot(data= cars, aes(x = speed ,y= dist)) +
  geom_point()
```

```{r, message=F}
ggplot(data= cars, aes(x = speed ,y= dist)) +
  geom_point() +
  geom_smooth()
```
```{r, message=F}
ggplot(data= cars, aes(x = speed ,y= dist)) +
  geom_point() +
  geom_smooth() +
  labs(title="What a nice relation!") +
  xlab("Car speed (km/h)") +
  ylab("Distance (km)")
```

*to be continued...*


